Dynamical Neural Network Models\\for Sequence Learning

### 2019-05-09 paper/main.tex
report

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[backend=biber,sorting=none]{biblatex}
%\usepackage{geometry}
\usepackage{float}
\addbibresource{mybib.tex}

@@

Cui et al. (2016) propose an artificial dataset of sequences of varying length (formed with characters) with overlapping subsequences. A sequence is sampled from the dataset and presented to the model at each time step. After the last character of the sequence is presented, a character is sampled from the noise distribution and shown to the model. This process is repeated until $10,000$ characters (counting both sequences and noise characters) are shown to the model. After the model sees $10,000$ characters, the last character of sequences with shared subsequences is swapped. The performance on this task is the accuracy of predicting the last character over a window of the last $100$ sequences. A sample dataset is summarized in Table \ref{tab:dataset}.

@@

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\linewidth]{../notebooks/matrix-distances.png}
    \caption{Heatmap of squared errors between distributed random encodings for the sequence characters, with $-1$ as the separating bit for ease of visualization.}
    \label{fig:encoding-distance}
\end{figure}

@@

\section*{Experiments}

\subsection*{Prediction Accuracy}

We tested each of the models (LSTM, TDNN, TDSNN, CSNN) on the discrete sequence learning task. The prediction accuracy results are shown in Figure \ref{fig:prediction-accuracy}. The prediction accuracy is the percent of correct predictions over a window of the last $100$ sequences.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\linewidth]{../results/artificial.png}
    \caption{Each of the models were tested on their ability to predict the last element of the sequence.}
    \label{fig:prediction-accuracy}
\end{figure}

The KNN (green) reaches perfect accuracy before the dataset is endings are swapped, but fails to relearn the data and only reaches $50\%$ accuracy afterward. The TDSNN (orange) is able to match the TDNN (teal) Surprisingly, the CSNN (blue) performance is almost able to match the LSTM (purple) performance. However the variance in the CSNN performance is much greater than the LSTM.

%\subsection*{Perturbation}

%Another important property of each of these models is the robustness to temporal noise. We introduce an $\alpha$ parameters which represents the probability that a character is swapped for another character.

%\subsection*{Continuous Dataset}

@@

We have shown two possible spiking architectures for learning changing patterns in time-series data in a semi-supervised fashion.

\newpage
\printbibliography[title={References}]
\end{document}

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[backend=biber,sorting=none]{biblatex}
%\usepackage{geometry}
\usepackage{hyperref}
\usepackage{float}
\addbibresource{mybib.tex}

@@

Cui et al. (2016) propose an artificial dataset of sequences of varying length (formed with characters) with overlapping subsequences. A sequence is sampled from the dataset and presented to the model at each time step. After the last character of the sequence is presented, a character is sampled from the noise distribution and shown to the model. This process is repeated until $10,000$ characters (counting both sequences and noise characters) are shown to the model. After the model sees $10,000$ characters, the last character of sequences with shared subsequences is swapped. The performance on this task is the accuracy of predicting the last character over a window of the last $100$ sequences. A sample dataset is summarized in Table \ref{tab:dataset}.  The code and datasets are open-source and available at \href{https://github.com/sharath/sequence-learning}{https://github.com/sharath/sequence-learning}.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.7\linewidth]{../notebooks/matrix-distances.png}
    \caption{Heatmap of squared errors between distributed random encodings for the sequence characters, with $-1$ as the separating bit for ease of visualization.}
    \label{fig:encoding-distance}
\end{figure}

@@

@@

\section*{Results}

\subsection*{Supervised Models}

\begin{figure}[H]
    \centering
    \includegraphics[width=\linewidth]{../results/artificial-supervised.png}
    \caption{The TDNN (teal) and TDSNN (orange) models are able to perfectly predict the sequence endings after $4,000$ elements, but their performance drops dramatically after the sequence endings are swapped. The LSTM (purple) is able to reach a reasonable performance on this task before the task and the drop after the sequences endings are swapped is only about $30\%$.}
    \label{fig:prediction-accuracy1}
\end{figure}

\subsection*{Semi-supervised Models}

\begin{figure}[H]
    \centering
    \includegraphics[width=\linewidth]{../results/artificial-unsupervised.png}
    \caption{The KNN (green) reaches perfect accuracy before the dataset is endings are swapped, but fails to relearn the data and only reaches $50\%$ accuracy afterward. The CSNN never manages to reach perfect accuracy on the task, but it is able to re-learn the sequence endings after they are swapped after experiencing a drop in performance similar to the LSTM in Figure \ref{fig:prediction-accuracy1}.}
    \label{fig:prediction-accuracy2}
\end{figure}

@@

\begin{table}[H]
    \centering
    \begin{tabular}{|c|c|} \hline
        Model & Average Accuracy \\ \hline
        LSTM  & $0.8049 \pm 0.088$ \\ \hline
        TDNN  & $0.8112 \pm 0.012$ \\ \hline
        TDSNN & $0.8097 \pm 0.013$ \\ \hline
        KNN   & $0.6391 \pm 0.021$ \\ \hline
        CSNN  & $\textbf{0.8250} \pm 0.119$ \\ \hline
    \end{tabular}
    \caption{The average accuracy on the discrete sequence learning task over $10$ randomized runs.}
    \label{tab:meanaccuracy}
\end{table}

We have shown two possible spiking architectures for learning changing patterns in time-series data. We tested each of the models (LSTM, TDNN, TDSNN, KNN, CSNN) on the discrete sequence learning task.  The TDSNN (orange) is able to match the TDNN (teal) Surprisingly, the CSNN (blue) performance over $10$ runs has the highest average accuracy over the sequence learning task, but it also has the average standard deviation over the $10$ runs.

\printbibliography[title={References}]
\end{document}
